Experience Report: Haskell in Computational Biology

### 2012-06-30 mrfy_experience_report.tex
fix line count

When performance has mattered, members of our group, like other
computational biologists, have used~C++.
To~compare our Haskell experience with our C++ experience,
we discuss three tools:
\begin{itemize}
\item
Matt~\cite{Menke:2008wu} is used to create alignments like that shown
in \figref{alignment}.
It~comprises about 12,000 lines of~C++.
The only external tools or libraries it uses are \texttt{zlib}
and \texttt{OpenMP}, and its initial development took two years.
\item
SMURF
\cite{Menke:2010ti} is used to detect homologous proteins in the presence
of paired beta strands.
It~comprises about 9,000 lines of~C++, of which about 1,000~lines are
shared with~Matt.
It~uses no external tools or libraries, and
its initial development took a year and a half.
It uses multidimensional dynamic programming to exactly compute the alignments
for which MRFy relies on stochastic search.
As a result, in the presence of complex beta-strand topologies, SMURF is
computationally intractable.
\item
MRFy is used to detect homologous proteins in the presence of paired
beta strands; it effectively supplants SMURF.
It~comprises about 1,200 lines of~Haskell.
\remark{HA HA HA.  It's over 3,500 lines now. Maybe around 800 lines
are tests or quickcheck properties?}
It~uses several external tools and libraries, of which the most
notable are Parsec, the BioHaskell bioinformatics library, and
the libraries \path{Data.MemoCombinators}, \path{Control.Parallel},
and \path{Data.Vector}.
Its initial development took about three months.
%%%% \remark{Others were mostly full-time work, but no need to distinguish.}
\end{itemize}
Like much research software, all three of these tools were written
 in~haste.
We~have experience modifying the older tools.

>>>>

When performance has mattered, members of our group, like other
computational biologists, have used~C++.
To~compare our Haskell experience with our C++ experience,
we discuss three tools:
\begin{itemize}
\item
Matt~\cite{Menke:2008wu} is used to create alignments like that shown
in \figref{alignment}.
It~comprises about 12,000 lines of~C++.
The only external tools or libraries it uses are \texttt{zlib}
and \texttt{OpenMP}, and its initial development took two years.
\item
SMURF
\cite{Menke:2010ti} is used to detect homologous proteins in the presence
of paired beta strands.
It~comprises about 9,000 lines of~C++, of which about 1,000~lines are
shared with~Matt.
It~uses no external tools or libraries, and
its initial development took a year and a half.
It uses multidimensional dynamic programming to exactly compute the alignments
for which MRFy relies on stochastic search.
As a result, in the presence of complex beta-strand topologies, SMURF is
computationally intractable.
\item
MRFy is used to detect homologous proteins in the presence of paired
beta strands; it effectively supplants SMURF.
It~comprises about 2300 lines of~Haskell, 300 of which are devoted
to tests and QuickCheck properties and generators.
Neither C++ codebase includes any testing facilities.
MRFy~uses several external tools and libraries, of which the most
notable are Parsec, the BioHaskell bioinformatics library, and
the libraries \path{Data.MemoCombinators}, \path{Control.Parallel},
and \path{Data.Vector}.
Its initial development took about three months.
%%%% \remark{Others were mostly full-time work, but no need to distinguish.}
\end{itemize}
Like much research software, all three of these tools were written
 in~haste.
We~have experience modifying the older tools.
